\bf Range Analysis for the IonMonkey JavaScript engine

### 2012-04-30 project-writings/paper.tex
Lots more paper work.

Thus we developed a quite simple and in practice efficient algorithm to
add $\beta$ nodes to an SSA source program by taking
advantage of data structures common to modern SSA based compilers.
\footnote{The method for removing $\beta$ nodes is so trivial that we
  could not claim with a straight face that we ``developed an
  algorithm'' to perform it.}

@@

function isDominatedUse(x', use):
  ins = use.instruction()
  if ins.isPhiNode():
    # Get the predecessor block that corresponds to the
    # argument of the phi node we are looking at
    phi-arg-pred-block = ins.block().pred(use.index())
    return blockDominates(x'.block(), phi-arg-pred-block)
  else:
    return blockDominates(x'.block(), x.block())
\end{verbatim}
\end{center}
\caption{Algorithm for inserting beta nodes}
\label{fig:add_betas}
\end{figure}

>>>>

Thus we developed a quite simple and in practice efficient algorithm to
add $\beta$ nodes to an SSA source program by taking
advantage of data structures common to modern SSA based compilers
\footnote{The method for removing $\beta$ nodes is so trivial that we
  could not claim with a straight face that we ``developed an
  algorithm'' to perform it.} .  Pseudocode for the algorithm is shown
in Figure \ref{fig:add_betas}. The algorithm traverses the control
flow group, looking for blocks that have been entered by a jump
conditioned on a relational operation. When we find a variable being
compared to a constant, we compute a range and use it to create a new
$\beta$ node. We want to propagate this range information to all uses
of the variable where the branch is guarenteed to have been taken;
this is accomplished by replacing all uses of original instruction
that are dominated by the new $\beta$ instruction with the $\beta$. To
do this, we take advantage of IonMonkey's use chains to efficiently
iterate over all uses of an instruction.

In order to do this, we needed to be able to efficiently determine
whether one block is dominated by another. The best way to do that
that existed in IonMonkey was to walk up the dominator tree looking
for the block being tested against. We felt that it would be too
inefficient to need to do this for every use of every variable tested
with a conditional. To do this, we annotated each basic block with its
index in a pre-order traversal of the dominator tree. Since IonMonkey
basic blocks already tracked how many other blocks they dominated, we
can now easily check whether a block dominates another by checking
whether the potential descendent's index falls within the range of the
ancestor's dominated blocks.

@@

function isDominatedUse(x', use):
  ins = use.instruction()
  if ins.isPhiNode():
    # Get the predecessor block that corresponds to the
    # argument of the phi node we are looking at
    phi-arg-pred-block = ins.block().pred(use.index())
    return blockDominates(x'.block(), phi-arg-pred-block)
  else:
    return blockDominates(x'.block(), x.block())

function blockDominates(b1, b2):
  high = b1.domIndex() + b1.numDominated()
  low  = b1.domIndex()
  return b2.domIndex() >= low && b2.domIndex <= high
\end{verbatim}
\end{center}
\caption{Algorithm for inserting beta nodes}
\label{fig:add_betas}
\end{figure}
